---
title: "`hyperSpec` download statistics"
author: "Claudia Beleites"
date:  "`r format(Sys.time(), '%d.%m.%Y')`"
output: html_document
---
# RStudio CRAN mirror

Aggregated server logs of the [RStudio CRAN mirror](http://cran-logs.rstudio.com) are publicly available. The provided data is likely not really representative for R users as I expect that only package installations/updates via the RStudio IDEs are covered. However, logs of other CRAN mirrors are AFAIK not publicly available.

```{r knitinit, echo = FALSE, results = 'hide', message = FALSE}
require ("knitr")
opts_chunk$set (echo = FALSE, results = 'hide', message = FALSE, warning = FALSE,
                dev='png', fig.width = 8, fig.height = 4.25, dpi = 100)
```
```{r}
start_date <- as.Date("2012-10-01")
yesterday <- as.Date (Sys.Date()) - 1
```
```{r load}
require ("dplyr")
require ("magrittr")

load ("logs.RData")
```
```{r download, echo=FALSE, eval = FALSE}
for (d in seq (start_date, yesterday, by = 1)){
  class (d) <- "Date"
  
  cat (format (d), "-------------------------------------------------------------------------------\n")
    
  if (! file.exists (format (d, "logs/%Y-%m-%d.csv.gz"))) {
    system (format (d, "wget -O logs/%Y-%m-%d.csv.gz http://cran-logs.rstudio.com/%Y/%Y-%m-%d.csv.gz"))
    system (format (d, "gunzip -kc logs/%Y-%m-%d.csv.gz > log"))
            
    system ("head -1 log > log.csv")
    system ("egrep hyperSpec log >> log.csv")
    
    logs <- rbind (logs, read.csv("log.csv"))
    
    save (logs, file = "logs.RData", compress = FALSE)
    unlink ("log")
    unlink ("log.csv")
  }
}

```

```{r}
require ("ggplot2")
require ("arrayhelpers")
require ("plyr")

summary (logs)

levels (logs$r_os)[grep ("darwin", levels (logs$r_os))] <- "Mac"
levels (logs$r_os)[grep ("linux-gnu", levels (logs$r_os))] <- "Linux"
levels (logs$r_os)[grep ("bsd", levels (logs$r_os))] <- "Linux"
levels (logs$r_os)[grep ("mingw", levels (logs$r_os))] <- "Win"
```

### Total downloads per month

```{r dl_month}
logs$month <- format (as.Date (as.character (logs$date)), "%Y-%m-01")
logs$month <- as.Date (as.character (logs$month))
ggplot (count (logs, "month"), aes (x = month, y = freq, group = 1)) + geom_bar(stat= "identity", width = 31) +
  #geom_point () + geom_line (size = 0.1) +
  ylim (0, NA) +
  ylab ("total no. of downloads via RStudio")
```


```{r}
monthly_os <- table(logs$month, logs$r_os, useNA = "always")
monthly_os <- monthly_os / rowSums (monthly_os)
monthly_os <- as.data.frame (monthly_os)
colnames (monthly_os) <- c ("month", "OS", "fraction")
monthly_os$month <- as.Date (as.character (monthly_os$month))
monthly_os$OS <- as.character(monthly_os$OS)
monthly_os$OS [is.na (monthly_os$OS)] <- "NA"
monthly_os$OS <- factor (monthly_os$OS, levels = c ("Linux", "Mac", "Win", "NA"))
monthly_os <- monthly_os [! is.na (monthly_os$month), ]
```


## Download statistics per country

```{r dl_country}
library(maptools)
library(mapproj)
library(rgeos)
library(rgdal)
library(ggplot2)
library(jsonlite)
library(RCurl)
require ("grid")


# naturalearth world map geojson
world <- readOGR (dsn = "ne_50m_admin_0_countries.geojson", layer = "OGRGeoJSON")

# remove antarctica
world <- world[!world$iso_a3 %in% c("ATA"),]
world <- spTransform(world, CRS("+proj=wintri"))

ccount <- count (logs, "country")
ccount <- ccount [order (-ccount$freq),]
rownames (ccount) <- NULL

## population data
## http://api.worldbank.org/v2/en/indicator/sp.pop.totl?downloadformat=csv
pop <- read.csv("sp.pop.totl_Indicator_en_csv_v2.csv", skip = 2, check.names = FALSE)
pop <- pop [, c ("Country Name", "Country Code", "2013"), ]
pop <- merge (pop, y = world@data [, c ("iso_a2", "iso_a3")], by.x = "Country Code", by.y = "iso_a3")
ccount <- merge (ccount, pop, by.x = "country", by.y = "iso_a2", all.x = TRUE)
ccount <- ccount [order (ccount$freq, decreasing = TRUE),]
ccount$dl.per.1e5capita <- ccount$freq / ccount$`2013` * 1e5

map <- fortify(world, region="iso_a2")

gg <- ggplot()
gg <- gg + geom_map (data=map, map=map, aes(x = long, y = lat, map_id = id, group = group), fill = "#ffffff", color = NA) 
gg <- gg + geom_map (data=ccount, map=map, color="white", size=0.15, aes (fill = freq, group = country, map_id = country))  
gg <- gg + scale_fill_gradientn(name="Downloads", colours = c ("#ffffa0", "darkgreen", rep ("#004000", 5)))
gg <- gg + coord_equal()
gg <- gg + theme(axis.line=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), 
                 axis.title=element_blank(), panel.background=element_rect(fill = "#e0e0ff"),
                 panel.border=element_blank(), panel.grid=element_blank(), panel.margin=unit(0, "lines"),
                 plot.background=element_blank(), legend.justification = c(0,0), legend.position = c (0, 0),
                 legend.key.size = unit (1.5, "cm"), legend.key.width = unit (0.5, "cm")) 
print (gg)
```

Here are the download numbers for the top 15 countries (since start of the server logs in October 2012):

```{r, results='markup'}
ccount [1:15, c ("country", "Country Name", "freq", "dl.per.1e5capita")]
```


And the downloads per population:
```{r}
gg <- ggplot()
gg <- gg + geom_map (data=map, map=map, aes(x = long, y = lat, map_id = id, group = group), fill = "#ffffff", color = NA) 
gg <- gg + geom_map (data=ccount, map=map, color="white", size=0.15, aes (fill = dl.per.1e5capita, group = country, map_id = country))  
gg <- gg + scale_fill_gradientn(name="Downloads\nper 100 000\ninhabitants", colours = c ("#ffffb0", "blue", "darkblue", "#000040", "#000040", "#000040"))
gg <- gg + coord_equal()
gg <- gg + theme(axis.line=element_blank(), axis.text=element_blank(), axis.ticks=element_blank(), 
                 axis.title=element_blank(), panel.background=element_rect(fill = "#e0e0ff"),
                 panel.border=element_blank(), panel.grid=element_blank(), panel.margin=unit(0, "lines"),
                 plot.background=element_blank(), legend.justification = c(0,0), legend.position = c (0, 0),
                 legend.key.size = unit (1.5, "cm"), legend.key.width = unit (0.5, "cm")) 
print (gg)
```


... and the top 15 countries in terms of downloads per population:
```{r, results='markup'}
ccount [order (-ccount$dl.per.1e5capita)[1:15], c ("country", "Country Name", "freq", "dl.per.1e5capita")]
```

### Version statistics

```{r}
vdf <- as.data.frame (table (logs$month, logs$version))
colnames (vdf) <- c ("month", "version", "freq")
vdf$month <- as.Date (as.character (vdf$month))
vdf$version <- factor (vdf$version, levels = sort (levels (vdf$version)))

ggplot (vdf, aes (x = month, y = freq, col = version, group = version)) + geom_point () + geom_line () 
```

### Operating System statistics
```{r}
ggplot (logs) + geom_bar(aes (x = r_os, fill = r_os)) +
  scale_fill_manual (values = c(Linux = "red", Mac = "darkblue", Win = "darkgreen"), na.value = "#808080") +
  theme (legend.position = "none")
```

I expect that the OS-dependent data may be quite distorted: I'd expect that installations on servers a) more frequently use Linux/Unix systems which b) install/update via closeby CRAN mirrors rather than the RStudio CRAN mirror. On the other hand, download statistics anyways underestimate the number of users for server systems.

```{r}
ggplot (monthly_os, aes (x = month, y = fraction, col = OS)) + 
  geom_point () + geom_line () +
  ylim (0, 1) +
  scale_color_manual (values = c(Linux = "red", Mac = "darkblue", Win = "darkgreen", `NA` = "#808080"))
```

In any case, the vast majority of `hyperSpec` users downloading via the RStudio CRAN mirror works on Windows.